---
title: "MOSAIC on scNMT_seq Mouse Gastrulation data"
output: github_document
---

```{r setup, include=FALSE}

library(MOSAIC)
library(pheatmap)
library(RColorBrewer)
library(genefilter)
library(knitr)
library(magrittr)
library(utilar)
library(kableExtra)
library(ggsci)
library(aricode)

knitr::opts_chunk$set(message = FALSE, echo=FALSE, collapse=TRUE, fig.retina=2, fig.path = "README_figures/README-")

```

## Introduction
We will be analyzing  [scNMT-seq](https://www.nature.com/articles/s41467-018-03149-4) study via MOSAIC to understand mouse gastrulation on their epigenome and transcriptome profiles to identify multi-omics signatures that characterize stage and lineage. 

[MOSAIC](https://github.com/arorarshi/MOSAIC) or Multi-Omics Supervised Integrative Clustering is a response weighted clustering algorithm inspired by [survClust](https://github.com/arorarshi/survClust), to classify samples into clusters that are relevant to outcome of interest.^1^ 

Each feature in a data type is weighed according to its association with binary or categorical outcome of interest, and a weighted distance matrix is computed ^2^. This reduces the computation space considerably from sample x feature to sample x sample. Samples are then projected into a multi dimensional space preserving the distance between them, and clustered with k-means algorithm to obtain class labels corresponding to outcome.

## Analysis 

```{r, echo=F, message=FALSE, results='hide'}

lapply(c("data/stage.Rdata", "data/lineage.all.Rdata", "data/lineage.Rdata", "data/cv.stats.noimp.github.Rdata"),load,.GlobalEnv)
```

We ran MOSAIC for __50 rounds__ of __5-fold__ cross validation for __k=2-7__, with stage and lineage as outcome of interest over 13 data types -

 * RNA
 * met_CTCF, met_DHS, met_p300, met_genebody, met_promoter, met_cgi
 * acc_CTCF, acc_DHS, acc_p300, acc_genebody, acc_promoter, acc_cgi 

and integrating all of them, to mine features that are associated with outcome of interest. 

RNA data was standardized, whereas proportion data from other data types was first transformed by taking their folded square root before standardizing. 

All the data was considered, including missing-ness, as MOSAIC can handle incomplete information among features and data types. If a data type had more than 5000 features, the feature space was reduce to top 5000 most variable features.  

Let's take a look at stage and lineage distribution. 

```{r, echo=F}

kable(get.summary(stage, 1, var.n="stage"), "html", caption="sample summary", row.names=TRUE )

kable(table(stage[names(lineage.all)], lineage.all), "html", caption="stage vs lineage", row.names=TRUE )

```

## Results 

We analyze MOSAIC obtained cross validated solutions over two metrics - adjusted Mutual Information (AMI) and Standardized Pooled Within Sum of Squares (SPWSS)

### Stage
```{r, fig.height=5, fig.width=11}
all.rdata<-c("acc_DHS", "acc_p300", "acc_cgi", "acc_CTCF", "acc_promoter", "acc_genebody", "met_DHS", "met_p300", "met_cgi", "met_CTCF", "met_promoter", "met_genebody", "rna.mat", "integrated")

plat.col = c("blue", "purple", "lightblue","darkblue","cyan", "cadetblue", "red", "darkred", "orange","coral","hotpink","brown", "darkolivegreen","black")

par(mfrow=c(1,2))
for (i in 1:length(all.rdata)){

  if(i==1){plot(x = c(2:7), y=apply(cv.stats[[i]]$AMI, 2, median), col=plat.col[i],type="o", xlab="k cluster", ylab="adjusted MI", bty="l", lwd=1.5, ylim=c(0,1), cex.lab=1.5, cex.axis=1.2)}
  
  if(i!=1){lines(x = c(2:7), y=apply(cv.stats[[i]]$AMI, 2, median), col=plat.col[i], type="o",lwd=1.5)}
  
}

legend("topright", all.rdata, col=plat.col, bty="n", lty=1, lwd = 2,ncol = 2, cex=0.9)

for (i in 1:length(all.rdata)){

  if(i==1){plot(x = c(2:7), y=apply(cv.stats[[i]]$spwss, 2, median), col=plat.col[i],type="o", xlab="k cluster", ylab="SPWSS", bty="l", lwd=1.5, ylim=c(0,1), cex.lab=1.5, cex.axis=1.2)}
  
  if(i!=1){lines(x = c(2:7), y=apply(cv.stats[[i]]$spwss, 2, median), col=plat.col[i], type="o",lwd=1.5)}
  
}

```

#### Let's take a look at RNA MOSAIC solution

```{r, fig.height=8}

lineage.nona = na.omit(lineage.all)
lineage.col = brewer.pal(length(unique(lineage.all)), "Set3")
names(lineage.col) = names(table(na.omit(lineage.all)))
           
stage.col = brewer.pal(length(unique(stage)), "PuBuGn")
names(stage.col) = names(table(na.omit(stage)))
ann_colors = list(stage = stage.col, lineage = lineage.col)
  
lineage.all = lineage.all[names(stage)]
aa=data.frame(stage = stage, lineage=lineage.all)
rownames(aa) = names(stage)

cl.col = pal_nejm("default")(7)

load("data/rna.mat.Rdata")
load("data/solnk.noimp.github.Rdata")
kk = solnk[[13]] 
kk = kk[order(as.numeric(kk))]
kk = factor(kk, levels=c("1","2","3","4","5"))
mat = t(rna.mat[names(kk),])
fp = rowFtests(mat, as.factor(kk))$p.value
names(fp) = rownames(mat)
fp.adj = p.adjust(fp)
fsig = fp.adj[which(fp.adj < 0.05)]
fsig.rna = fsig[1:1000]
fsig = sort(fsig)[1:500]

#top 500 features
imagedata = mat[names(fsig),names(kk)]
imagedata[imagedata < -2] = -2
imagedata[imagedata > 2] = 2

xaa = aa[names(kk),]
xaa$cluster=kk

kk.col = cl.col[1:length(unique(kk))]
names(kk.col) = as.character(unique(kk))
xann_colors = ann_colors; xann_colors$cluster = kk.col

pheatmap(imagedata, clustering_method ="ward.D",scale='none',annotation_col=xaa, annotation_colors=xann_colors, cluster_cols = FALSE,show_colnames=F,show_rownames = FALSE, border_color =NA, fontsize_row = 8,main = "RNA 5-class MOSAIC vs stage, top500", color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100))

kkAMIstage = AMI(kk, stage[names(kk)])
kable(table(kk, stage[names(kk)]), "html", caption=paste0("MOSAIC rna vs stage, AMI= ",round(kkAMIstage,2)), row.names=TRUE )

kkAMIlin = AMI(kk[names(lineage.nona)], as.character(lineage.nona))

kable(table(kk, lineage.all[names(kk)]), "html", caption=paste0("MOSAIC rna vs lineage, AMI=",round(kkAMIlin,2)), row.names=TRUE )

```

__Let us take a look at some analysis with kmeans__

```{r}

set.seed(123)
unwt.dd<-as.matrix(dist(rna.mat))
cmd.mat = cmdscale(unwt.dd, nrow(unwt.dd)-1)
unwtkk = kmeans(cmd.mat, 5, nstart=100)
unwtkkAMI = AMI(unwtkk$cluster, stage)

unwtkkAMIstage = AMI(unwtkk$cluster, stage[names(unwtkk$cluster)])
kable(table(unwtkk$cluster, stage[names(unwtkk$cluster)]), "html", caption=paste0("kmeans rna vs stage, AMI= ",round(unwtkkAMIstage,2)), row.names=TRUE )

unwtkkAMIlin = AMI(unwtkk$cluster[names(lineage.nona)], as.character(lineage.nona))

kable(table(unwtkk$cluster, lineage.all[names(unwtkk$cluster)]), "html", caption=paste0("kmeans rna vs lineage, AMI=",round(unwtkkAMIlin,2)), row.names=TRUE )


```

#### Other MOSAIC solutions
Make a _circomap_ of remaining subtypes, and stage and lineage classification 

```{r, fig.height=8, fig.width=8}

library(panelmap)
names(solnk) = all.rdata
solnk.AMI = round(unlist(lapply(solnk, function(x) AMI(x, stage[names(x)]))),2)

tsolnk = solnk[1:12]
names.circo = paste0(names(solnk),";AMI=", solnk.AMI)

circo.dat<-list()
for(i in 1:length(solnk)){
  
  cl = solnk[[i]][order(as.numeric(solnk[[i]]))]
  stage.cl = stage[names(cl)]
  lineage.cl = lineage.all[names(cl)]
  
  tt = cbind(cl, stage.cl, lineage.cl)
  
  circo.dat[[i]] = tt  
  
}

names(circo.dat) = names.circo
gtoplot<-c("cl")
gcol = cl.col
gheight = 0.10

ftoplot<-c("stage.cl","lineage.cl")
ftype= c(1,1)
fcol=list(stage.cl = stage.col, lineage.cl= lineage.col)

fheight<-list(); fheight[1:length(ftoplot)] = 0.10
#Voila!!

circomap(circo.dat[1:6], gtoplot, gcol, gheight, ftoplot, ftype, fcol, fheight, NA.flag=TRUE)
circomap(circo.dat[7:12], gtoplot, gcol, gheight, ftoplot, ftype, fcol, fheight, NA.flag=TRUE)

```

### Integrated solution

```{r}

kk=solnk[[14]]
kkAMIstage = AMI(kk, stage[names(kk)])
kable(table(kk, stage[names(kk)]), "html", caption=paste0("integrated vs stage, AMI= ",round(kkAMIstage,2)), row.names=TRUE )

kkAMIlin = AMI(kk[names(lineage.nona)], as.character(lineage.nona))

kable(table(kk, lineage.all[names(kk)]), "html", caption=paste0("integrated vs lineage, AMI=",round(kkAMIlin,2)), row.names=TRUE )

rnakk = solnk[[13]]
kkAMIrnakk = AMI(kk, rnakk[names(kk)])

kable(table(kk, rnakk[names(kk)]), "html", caption=paste0("integrated vs rna, AMI=",round(kkAMIrnakk,2)), row.names=TRUE )

```

## Conclusion

* MOSAIC finds supervised clusters, with an outcome of interest in mind. Where kmeans might give mixed results. Supervised clustering is much more efficient and helps in sorting out different signals 

* MOSAIC can run with missing data. However interpretations should be made carefully.
 
* MOSAIC reduces computation space from sample x feature to sample x sample

* Efficient in dealing with noisy features

## References

1. Arora A, Olshen AB, Seshan VE, and Shen R. Pan-cancer identification of clinically relevant genomic subtypes using outcome-weighted integrative clustering. Biorxiv

2. Xing, E. P., Jordan, M. I., Russell, S. J., & Ng, A. Y. (2003). Distance metric learning with application to clustering with side-information. In Advances in neural information processing systems (pp. 521-528).