---
title: "MOSAIC on scNMT_seq Mouse Gastrulation data"
output: github_document
---

```{r setup, include=FALSE}

#load libraries and other utilities
source("../MOSAIC.R")

library(mclust)
library(pheatmap)
library(RColorBrewer)
library(genefilter)
library(knitr)

knitr::opts_chunk$set(collapse=TRUE, fig.retina=2, fig.path = "README_figures/README-", echo=FALSE, message = FALSE)

```

## Introduction
We will be analyzing  [scNMT-seq](https://www.nature.com/articles/s41467-018-03149-4) study via MOSAIC to understand mouse gastrulation on their epigenome and transcriptome profiles to identify multi-omics signatures that characterize stage and lineage. 

[MOSAIC](https://github.com/arorarshi/MOSAIC) or Multi-Omics Supervised Integrative Clustering is a response weighted clustering algorithm inspired by [survClust](https://github.com/arorarshi/survClust), to classify samples into clusters that are relevant to outcome of interest. 

Each feature in a data type is weighed according to its association with binary or categorical outcome of interest, and a weighted distance matrix is computed[1]. This reduces the computation space considerably from sample x feature to sample x sample. Samples are then projected into a multi dimensional space preserving the distance between them, and clustered with k-means algorithm to obtain class labels corresponding to outcome.

## Preliminary Analysis 

We ran MOSAIC for 10 rounds of 3-fold cross validation, with stage as outcome of interest over seven data type - RNA, met_CTCF, met_DHS, met_p300, acc_CTCF, acc_DHS, acc_p300, and integrating all of them, to mine features that are associated with stage. 

RNA data was standardized, whereas proportion data from other data types was first transformed by taking their folded square root before standardizing. 

All the data was considered, including missing-ness, as MOSAIC can handle incomplete information among features and data types. 

We analyze MOSAIC obtained cross validated solutions over two metrics - adjustedRandIndex and Standardized Pooled Within Sum of Squares (SPWSS)

```{r, fig.height=5, fig.width=11}
load("../Supervised_iCluster/MOSAIC/stage.Rdata")

plat = c("acc_DHS", "acc_CTCF", "acc_p300", "met_DHS", "met_CTCF", "met_p300", "rna.mat", "integ")
plat.col = c("blue", "purple", "lightblue", "red", "darkred", "orange", "darkolivegreen", "black")

get.stats<-function(cv.fit,truth, kk=c(2:7), cvr = 10){
  
  spwss = adjR = matrix(NA, nrow=cvr, ncol=length(kk))
  for(i in 1:length(cv.fit)){
    tt = cv.fit[[i]]
    spwss[,i] = unlist(lapply(tt, function(x) x$cv.ss))
    adjR[,i] = unlist(lapply(tt, function(x) adjustedRandIndex(x$cv.labels, truth[names(x$cv.labels)])))
    
  }
  return(list(spwss = spwss, adjR=adjR))
  
}

par(mfrow=c(1,2))
for (i in 1:length(plat)){
  load(paste0("../Supervised_iCluster/MOSAIC/",plat[i], "_cvfit.Rdata"))
  tt = get.stats(cv.fit, stage)
 
  if(i==1){plot(x = c(2:7), y=apply(tt$adjR, 2, median), col=plat.col[i],type="o", xlab="k cluster", ylab="adjusted Rand Index", bty="l", lwd=1.5, ylim=c(0,0.5), cex.lab=1.5, cex.axis=1.2)}
  
  if(i!=1){lines(x = c(2:7), y=apply(tt$adjR, 2, median), col=plat.col[i], type="o",lwd=1.5)}
  
}

legend("topright", plat, col=plat.col, bty="n", lty=1, lwd = 2,ncol = 3)

for (i in 1:length(plat)){
  load(paste0("../Supervised_iCluster/MOSAIC/",plat[i], "_cvfit.Rdata"))
  tt = get.stats(cv.fit, stage)
 
  if(i==1){plot(x = c(2:7), y=apply(tt$spwss, 2, median), col=plat.col[i],type="o", xlab="k cluster", ylab="SPWSS", bty="l", lwd=1.5, cex.lab=1.5, cex.axis=1.2, ylim=c(0,1))}
  
  if(i!=1){lines(x = c(2:7), y=apply(tt$spwss, 2, median), col=plat.col[i], type="o",lwd=1.5)}
  
}

```

We see that RNA expression predicts stage the best, at k=4 with maximum adjusted Rand Index, followed by integration and met_CTCF. 

let's take a look at RNA, k=4 solution. 

```{r}

load("../Supervised_iCluster/MOSAIC/rna.mat_cvfit.Rdata")
load("../Supervised_iCluster/MOSAIC/rna.mat.dd.Rdata")
rnak4 = cv.voting(cv.fit, rna.mat.dd,4)
kable(table(rnak4[names(stage)], stage), "html", row.names=T, caption="MOSAIC RNA 4-class vs stage")
nf=50
load("../Supervised_iCluster/MOSAIC/rna.mat.Rdata")
mat = rna.mat[names(rnak4), ]
f=rowFtests(t(mat),fac=as.factor(rnak4))$p.value
adj.P.Val=p.adjust(f)
if(is.null(nf)){nf=2000}
imagedata=mat[,order(adj.P.Val)[1:nf]]

imagedata[imagedata < -3] = -3
imagedata[imagedata > 3] = 3

Colors=rev(brewer.pal(11,"Spectral"))
Colors=colorRampPalette(Colors)(100)
oo = order(rnak4)

soln.col = brewer.pal(length(unique(rnak4)), "Dark2")
ann_colors = list(cluster = c("1"= soln.col[1],"2" =soln.col[2], "3" = soln.col[3], "4" = soln.col[4]))
cluster = rnak4[oo]
aa = data.frame(cluster)

pheatmap(t(imagedata[oo,]),scale='none',annotation_col=aa, annotation_colors=ann_colors, show_colnames=F, border_color =NA, fontsize_row = 8, cluster_cols=F,main = "MOSAIC 4-class RNA, top 50 features", color = Colors)

```



## Future Work

* Study the various MOSAIC solutions of different platforms. 
* Can integrated perform better? 
* Take into account other data types like `cgi`, `promoter` and `genebody`
* With lineage as outcome

## Reference

1. Xing, E. P., Jordan, M. I., Russell, S. J., & Ng, A. Y. (2003). Distance metric learning with application to clustering with side-information. In Advances in neural information processing systems (pp. 521-528).
